---
title: "Tutorial 2"
author: "Nicholas Kiriazis"
date: "2022-09-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
```

## Reading in the data

The data file can be found on kaggle: https://www.kaggle.com/datasets/dansbecker/nba-shot-logs?resource=download

```{r}
data_tibble = read_csv("tut02_09_16_data.csv")
```

```{r}
head(data_tibble)
```

Assume we are only interested in certain columns
```{r}
data_tibble = data_tibble[,c("LOCATION", "PERIOD", "SHOT_CLOCK", "DRIBBLES", "SHOT_DIST", "PTS_TYPE", "SHOT_RESULT", "player_name")]

data_tibble["SHOT_RESULT"] = (data_tibble["SHOT_RESULT"] == "made")

data_tibble
```

## Exploring the data

Quick reminder about displaying
```{r}
c(1,2,3)
```

Now we make some plots to see if the data is okay.
```{r}
two_tibble = data_tibble[data_tibble$PTS_TYPE == 2,]

two_plot = ggplot(two_tibble, aes(x=SHOT_DIST))
```

```{r}
two_plot = two_plot + geom_histogram(bins=30, col="black", fill="lightblue")
```

```{r}
print(two_plot)
```

```{r}
two_plot = two_plot + geom_vline(xintercept=24, color="red")
```

```{r}
print(two_plot)
```

## Filtering data

We need to remove the impossible two point shots
```{r}
clean_tibble = data_tibble %>% filter(!(SHOT_DIST > 24 & PTS_TYPE==2))
```

Let's just make sure we filtered okay.
```{r}
ggplot(
  filter(clean_tibble, PTS_TYPE==2), aes(x=SHOT_DIST)) + 
  geom_histogram(bins=30, col="black", fill="lightblue") +
  geom_vline(xintercept=24, color="red"
)
```
## Doing fancy analysis

Is there an advantage to shooting at home?
```{r}
loc_group = group_by(clean_tibble, LOCATION)

loc_percentage = summarise(loc_group, RATE=sum(SHOT_RESULT)/length(SHOT_RESULT))
loc_percentage
```

Does this trend hold for all players
```{r}
player_loc_percentage = 
  group_by(clean_tibble, LOCATION, player_name) %>%   
  summarise(RATE=sum(SHOT_RESULT)/length(SHOT_RESULT)) %>% 
  ungroup()

player_loc_percentage = arrange(player_loc_percentage, player_name)

player_loc_percentage
```
Let's re-organize this into a more usable format
```{r}
home_values = filter(player_loc_percentage, LOCATION=="H")
print(home_values)
away_values = filter(player_loc_percentage, LOCATION!="H")
print(away_values)
```
Since the values are in the same order across dataframes, we can easily work with this
```{r}
home_away_diff_tbl = tibble(HOME=home_values$RATE, AWAY=away_values$RATE)

ggplot(home_away_diff_tbl, aes(x=HOME, y=AWAY)) + geom_point() + geom_abline(intercept = 0, slope=1, col="red")
```