---
title: "Tutorial 1"
author: "Nicholas Kiriazis"
date: "2022-09-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NOTE: Some of the cells will lead to errors. This is for illustrative purposes.

A lot of cells have no description, the idea is that you should think about what the code will do, and run the cell AFTER to see the behavior.

## Reading Data

```{r}
test = read.csv("Tutorial01_09-09_data.csv")

test
```

```{r}
test = read.csv("Tutorial01_09-09_data.csv", colClasses = c("character", "numeric"))

test
```

## Vectors

### Declarations

```{r}
y = vector(length=2)
y[1] = "a"
y[2] = "b"

y
```

```{r}
y[3]
```

```{r}
y[3] = "c"
y
```


### Quirks of casting

Defining some test vectors.
```{r}
c1 = c(1,2,3)
c2 = c(4,5,6,0,0,0)
c3 = c(7,8)
c4 = c(1,2,"3")
```

Some quirky examples of casting 

```{r}
c1 + 7
```

```{r}
c1+c1
```

```{r}
c1 + c2
```

```{r}
c1*c3
```

```{r}
c1 + c4
```

```{r}
c1[1] + c4[1]
```

## Lists

Shamelessly stealing Professor Steele's list example
```{r}
U1 = c(203, 204)
U2 = c(323,324)
U3 = c(208,427,423,523,545)

mymcgill_stats = list(U1,U2,U3,"Statistics Major")

mymcgill_stats
```

Run this cell twice and see what happens
```{r}
print(names(mymcgill_stats))

names(mymcgill_stats) = c("U1", "U2", "U3", "Major")

print(names(mymcgill_stats))

print("------")
mymcgill_stats
```


Trying to access elements
```{r}
mymcgill_stats$Major
```


```{r}
mymcgill_stats[1] 
mymcgill_stats[2] 

mymcgill_stats[1] + mymcgill_stats[2]
```

```{r}
mymcgill_stats[1] 
mymcgill_stats[[1]] 
```

```{r}
mymcgill_stats[[1]] + mymcgill_stats[[2]]
mymcgill_stats$U1 + mymcgill_stats$U2
```


```{r}
mymcgill_stats[c(1,2)]
```

```{r}
mymcgill_stats[[c(1,2)]]
```
Often times "mistakes" will still produce output, just not in ways you expect. For example, assuming you want to look at the elements inside the U1 list.
```{r}
# This loops through the courses
u1_a = mymcgill_stats[[1]]

for(i in 1:length(u1_a)){
  print(u1_a[i])
}

print("-----")

# This loops through a list with one element
u1_b = mymcgill_stats[1]

for(i in 1:length(u1_b)){
  print(u1_b[i])
}
```

## Named vectors

```{r}
test_vector = c(1,2,3)
test_vector
names(test_vector) = c("a", "b", "c")
test_vector
```
The "class" function is infinitely useful (especially when working with a new package).
```{r}
class(mymcgill_stats)
```

```{r}
test_vector[1]
```

```{r}
test_vector[4] = 4

test_vector

```

```{r}
test_list = as.list(test_vector)
test_list
```

```{r}
test_list$a
```

