---
title: "Assignment1-Q1"
output: pdf_document
date: "2022-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2

a. Read the data into R using the read_csv function and save it to an R object named club_teams. Give the code you used to import the data into R and use the head() function to show the ﬁrst six rows of data.

```{r} 
library(tidyverse)
```

```{r}
club_teams = read_csv("spi_global_rankings.csv")
head(club_teams)
```

b. What class of object is club_teams?

```{r}
class(club_teams)
```

c. What is the mode of the column titled name?

```{r}
mode(club_teams$name)
```

d. Add a column (named: diffxG) to the club_teams object containing the diﬀerence between the oﬀensive expected goals (off) and the defensive expected goals (def).

```{r}
club_teams["diffxG"] = club_teams["off"] - club_teams["def"]
head(club_teams)
```

e. Using slice and filter, write a line (or a few lines) of code that will return the ﬁrst ﬁve rows from the league Italy Serie A and save it to an object called first_five_serieA. Give your code and print this object (using print) to show that you have done it correctly. Recall: the == operator can be used to test for whether the value on the left hand side in a vector is equal to the value no the right hand side. Hint: the correct order of slice and ﬁlter matters!

```{r}
first_five_serieA = slice(filter(club_teams, league == "Italy Serie A"), 1:5)
print(first_five_serieA)
```


f. Add a new column to club_teams called SPI_max_league_diff which contains the diﬀerence in SPI between the maximum SPI value **for that league* and the team in each row. In other words, the best team in each league should have an SPI_max_league_diff value of 0 and the second best team in each league should have a positive number (the diﬀerence between the league maximum and thir own SPI value). Hint: Think about the kid’s height example from class. Give your code for adding the column and use head to show the ﬁrst six rows of the updated object.

```{r}
club_teams <- club_teams %>% 
  group_by(league) %>%
  mutate(SPI_max_league_diff = max(spi)-spi) %>%
  ungroup()
head(club_teams)
```
